---
title: "plotly"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

Click [here](index.html) to return to the index page!

We're gonna look at the NYC Restaurant Inspections data.
```{r load data}
library(flexdashboard)
data("rest_inspec")
library(tidyverse)
library(p8105.datasets)
library(plotly)
```

```{r}
data("rest_inspec")

```

```{r}
rest_inspec <- rest_inspec %>%
  filter(!is.na(score) & 
         boro %in% c("BRONX", "MANHATTAN", "BROOKLYN") &
         cuisine_description %in% c("Mexican", "Chinese", "Japanese", "Italian", "Jewish/Kosher", "Russian", "American", "Middle Eastern", "Spanish", "Greek", "Thai")) %>%
  select(boro, cuisine_description, dba, critical_flag, score, grade)

```

```{r}
print(rest_inspec)
```

The `rest_inspec` dataset contains `r nrow(rest_inspec)` observations and ``r ncol(rest_inspec)`` variables.

We only want to look at restaurants located in the boroughs of "Manhattan", "BRONX", and "BROOKLYN".
Cuisine Filter: The dataset was filtered to include only specific cuisine types: "Mexican", "Chinese", "Japanese", "Italian", "Jewish/Kosher", "Russian", "American", "Middle Eastern", "Spanish", "Greek", and "Thai".
Only relevant columns were retained for analysis: cuisine_description, dba, critical_flag, score, and grade.

# 1. Which cuisine types are most common in each borough?
```{r most common in each borogh}
cuisine_counts <- rest_inspec %>%
  count(boro, cuisine_description) %>%
  arrange(boro, desc(n))

fig_1 <- plot_ly(data = cuisine_counts, x = ~boro, y = ~n, type = 'bar', color = ~cuisine_description, colors = 'viridis') %>%
  layout(yaxis = list(title = 'Number of Establishments'),
         xaxis = list(title = 'Borough'),
         barmode = 'stack',
         title = 'Count of Cuisine Types by Borough')
fig_1 
```
American cuisine is the most common in Brooklyn, Manhattan and bronx, follow by Chinese cuisine. 

# 2. How does the average health inspection score differ among the various cuisines?
```{r}
color_palette <- scales::hue_pal()(length(unique(rest_inspec$cuisine_description)))

fig_2 <- rest_inspec %>%
  plot_ly(x = ~cuisine_description, y = ~score, type = 'box', color = ~cuisine_description, colors = color_palette) %>%
  layout(yaxis = list(title = 'Health Inspection Scores'),
         xaxis = list(title = 'Cuisine Description'),
         title = 'Health Inspection Scores by Cuisine')
fig_2
```

# 3. What's the distribution of grades across the different boroughs?
```{r}
grade_distribution <- rest_inspec %>%
  count(boro, grade) %>%
  arrange(boro, grade)

fig_3 <- plot_ly(data = grade_distribution, x = ~boro, y = ~n, type = 'bar', color = ~grade, colors = 'Paired') %>%
  layout(yaxis = list(title = 'Count of Grades'),
         xaxis = list(title = 'Borough'),
         barmode = 'stack',
         title = 'Distribution of Grades Across Boroughs')
fig_3
```
The majority of restaurants loctaed in Bronx, Brooklyn and manhattan have grade A.





